Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-03-13 WSSSPE2_report/WSSSPE2_review.tex
adding initial prizes text - more to be done by Suresh

\subsubsection{Remaining breakout discussions: Prizes}\label{sec:prizes}
\note{writing in progress by Suresh Marru - in a GitHub branch to be merged?}

\url{http://tinyurl.com/mjok4o5}
%https://docs.google.com/document/d/1foSppeHQeQ-qOjd-Rp_CUCNy8JMLxd_vrCDMmqvG2-w/edit

>>>>

\subsubsection{Remaining breakout discussions: Prizes}\label{sec:prizes}
\note{sub-breakout notes: \url{http://tinyurl.com/mjok4o5}}
%https://docs.google.com/document/d/1foSppeHQeQ-qOjd-Rp_CUCNy8JMLxd_vrCDMmqvG2-w/edit

This group discussed the idea of prizes.  Prizes are expected to reframe software
as ``instrument building'' but will prizes be good or bad, and how can we make sure
there are no negative affects and the process cannot be gamed?

Prizes in different categories were discussed (like Academy Awards), for example:
best contribution (non-founder),
broadest diversity of contributions,
best tutorials/documentation,
best leadership transition (award ex leader and new leader),
best generalization (taking something that was limited and making it more general), and
best mentorship of contributors (bringing others into the community).

\katznote{I don't understand this at all.  Can some explain it?}
There are also other indirect forms of incentives
(though these are not a direct prize, they still can incentivize someone):
\begin{itemize}
\item Converting reputation by joining apache software foundation, Google etc. 
\item Inviting ASF and open source people to contribute to scientific code (at least on questions of overlapping
\item Template for assessing scientific contributions made through software.   
\end{itemize}

To determine who should give prizes, perhaps we should find those who we think would
be awarded prizes, then ask them who they would want to receive a prize from, then
ask those organizations to see if they are willing to be involved in the process.

One idea is to create a funding program for disciplines or other organizations to
create a prize program.  We provide a framework, for example: awards to individuals, must
award in 5 or 6 areas, must have a jury and/or objective criteria that includes
senior/junior domain people and technology people, must have the recipients awarded
at a relevant event, must provide citations that explain why the prizes should be awarded.
A concern is that having many organizations award prizes may reduce the impact.

Possible awarders are AAAS, Nature Publishing Group, ACM/IEEE, Astro, Ecology Society of America, etc.  Or perhaps this should be a joint technology/science partnership, for example, the [Apache $|$ Mozilla]--[AAAS $|$ disciplinary society] prize?

Some potential criteria for prizes are:
community engagement, helping out others;
number of unique contributors;
adding new pieces of functionality to software;
integrating software into broader ecosystem; championing broad principles of sustainability, open science, open source, etc.;
improving accessibility to software, to scientific software (perhaps championing inclusiveness or making software accessible?);
documentation;
tutorials;
commits / patching;
leadership transition; and
best contribution by a non-founder.

Perhaps there should be different criteria for ``established'' members of the community versus junior members?  Or perhaps prizes should be restricted to junior members?

An important point is that we don't want to give prizes just to reward people who are really good at this one thing, but rather we want to reward people who are building the culture we want as scientists.
